Agile Monitoring of Evolving Web Services: For the Little Man

### 2018-04-06 grouping.tex

\begin{itemize}
	\item \code{'User'} stands for the name of this particulart grouping strategy
	\item \code{lambda: Session.find(... } is a callback with no arguments, 
	that makes use of the global flask.request and extracts from 
	the request the current session, and in turn the user id\footnote{
		Note that every web service or application must have a way of 
		associating a request with a user. In fact, the Session.find(request) 
		was already an existing function in the analyzed system}
\end{itemize}

In our case study, we wanted to understand the per-user performance of \epFeedItems 
--  an endpoint that retrieves a list of recommended articles for a given user.
It is the endpoint with the slowest response time and highest variability (see \Fref{fig:ep}).

% Once the \tool is configured with the line of code aforementioned.

\begin{figure}[h!]
  \centering
  \includegraphics[width=.5\linewidth]{time_per_user}
  \caption{The \epFeedItems shows a very high variability across users}
  \label{fig:tpu}
\end{figure}

>>>>

\begin{itemize}
	\item \code{'User'} stands for the name of this particular grouping strategy
	\item \code{lambda: Session.find... } is a callback with no arguments, 
	that makes use of the global \code{flask.request} and extracts from 
	the request the current session, and in turn the user id\footnote{
		Note that every web service or application must have a way of 
		associating a request with a user. In fact, the Session.find(request) 
		was already an existing function in the analyzed system}
\end{itemize}

In Zeeguu, it was decided to further understand the per-user performance of \epFeedItems
--  an endpoint that retrieves a list of recommended articles for a given user.
It is the endpoint with the slowest response time and highest variability (see \Fref{fig:ep}).

\begin{figure}[h!]
  \centering
  \includegraphics[width=.7\linewidth]{time_per_user}
  \caption{The \epFeedItems shows a very high variability across users}
  \label{fig:tpu}
\end{figure}

### 2018-04-06 paper.tex
intro to evolution

\author{
\IEEEauthorblockN{The Author, The Other Author}\\
Institute\\
University, Country\\
Email: \{author\}@university.domain 
}

@@

\input{architecture}

>>>>

\author{
\IEEEauthorblockN{The Author, The Other Author}\\
Johann Bernoulli Institute for Mathematics and Computer Science\\
University of Groningen, Groningen, the Netherlands\\
Email: \{author\}@university.domain 
}

@@

% \input{architecture}

### 2018-04-06 testing.tex
cleaning up stuff
...

\newpage
  \section{Performance Regression Monitoring}

@@

  Each time a new build is created through Travis, the \tool automatically detects all available unit tests defined by the application developer (in (-{}-test\_folder)) and iterates through each one of them a number of times (-{}-times) while monitoring the response times for each test. The resulting measurements are uploaded to a specific endpoint that is created by the tool at the API url (-{}-url). The data uploaded to this endpoint is   persisted in a separate part of the tool database so as not to contaminate the ``live'' monitoring data.

      \begin{figure}[h!]
        \centering
        \includegraphics[width=0.45\columnwidth]{travis_builds}
        \caption{Response times in 5 consequent Travis builds}
        \label{fig:builds}
      \end{figure}

  \Fref{fig:builds} is a screenshot from the dashboard showing the measured response times for 5 consequent builds, with 180 iterations of the unit tests executed in total across all endpoints. The outliers on the right of the figure are due to initial requests \ins{which must wait for a boot-up phase of the API}.

@@

  While this integration testing environment is different from the production one, and the load used is purely synthetic, it can serve as an early performance indicator for the developer.  Figure \ref{fig:response_times_preemptive} shows the actual endpoint response times across five versions of the system (above) and the corresponding testing times for the synthetic load in the same five versions (below). The green line joining the medians hints to a correlation between the two sets of measurements.

@@

  Computing Pearson correlation ($r(3)=.93, p=.02$) between the median values of the two datasets (cf. Table \ref{tab:correlations}) shows that there is a correlation between the two.

@@

  \ml{This following paragraph is beautifully written, but it's a bit redundant with what it's already said. However maybe after some more refactorings we can put it to use. Worst case, we move it to the conclusion}
  \ins{Besides functioning as an early warning system for performance, tracking the evolving performance of API tests serves a second purpose. To function as an anchor when the developer analyzes performance degradations. For example, when a developer sees that a given endpoint has become less performant after in a newly deployed version, they are able to investigate whether the performance degradation is visible also with the synthetic load. This would correspond to a performance degradation which is due to ``algorithmic performance degradation''. If on the other hand, the performance of the tests does not change between versions, the developer might conclude that the performance degradation could be due to the workload on the machines, or maybe to the workload mix of the users}

>>>>

  \section{Performance Triangulation With Regression Monitoring}
  \label{sec:regression}

@@

  Each time a new build is created through Travis, the \tool automatically detects all available unit tests defined by the application developer (in (\code{-{}-test\_folder)}) and iterates through each one of them a number of times (\code{-{}-times}) while monitoring the response times for each test. The resulting measurements are uploaded to a specific endpoint that is created by the tool at the API url (\code{-{}-url}). The data uploaded to this endpoint is   persisted in a separate part of the tool database so as not to contaminate the ``live'' monitoring data.

      \begin{figure}[h!]
        \centering
        \includegraphics[width=0.65\columnwidth]{travis_builds}
        \caption{Response times in 5 consequent Travis builds}
        \label{fig:builds}
      \end{figure}

  \Fref{fig:builds} is a screenshot from the dashboard showing the measured response times for 5 consequent builds, with 180 iterations of the unit tests executed in total across all endpoints. The outliers on the right of the figure are due to initial requests \ins{which must wait for a boot-up phase of the API}. \va{Remember to make greyscale for consistency}

@@

  While this integration testing environment is different from the production one, and the load used is purely synthetic, it can serve as an early performance indicator for the developer.  \Fref{fig:response_times_preemptive} shows the actual endpoint response times across five versions of the system (above) and the corresponding testing times for the synthetic load in the same five versions (below). The green line joining the medians hints to a correlation between the two sets of measurements.

@@

  Computing Pearson correlation ($r(3)=.93, p=.02$) between the median values of the two datasets (cf. Table \ref{tab:correlations}) shows that there is indeed a correlation between the two.

@@

  \ml{This following paragraph is beautifully written, but it's a bit redundant with what it's already said. However maybe after some more refactorings we can put it to use. Worst case, we move it to the conclusion}
  \va{I concur}
  \ins{Besides functioning as an early warning system for performance, tracking the evolving performance of API tests serves a second purpose. To function as an anchor when the developer analyzes performance degradations. For example, when a developer sees that a given endpoint has become less performant after in a newly deployed version, they are able to investigate whether the performance degradation is visible also with the synthetic load. This would correspond to a performance degradation which is due to ``algorithmic performance degradation''. If on the other hand, the performance of the tests does not change between versions, the developer might conclude that the performance degradation could be due to the workload on the machines, or maybe to the workload mix of the users}

### 2018-04-06 utilization.tex

  The default landing page of the dashboard presents an overview of all the endpoints, 
  overview which presents some basic utilization information as:
  (1) the number of calls to that endpoint since the begining of tracking
  (2) the number of calls in the last seven days, and
  (3) the number of calls in the last day\footnote{Although very useful, it was only after several months of using the system that the client requested the extra columsn for one day and seven days. An argument for slow research}s

@@

        \item \epTranslationsColor is an indicator of the amount of foreign language reading the users are doing

@@

  For a more detailed evolution of utilization, Figure \ref{fig:aeu} shows the \perspective{Daily API Utilization} perspective on API utilization that \tool provides: a stacked bar chart of the number of hits to various endpoints grouped by day\footnote{Endpoint colors are the same in different views}. Figure~\ref{fig:aeu} in particular shows a peak utilization, on June 28 when the API had more than 12.000 hits\footnote{Turns out that the highschool students that were using the system for their class had a deadline on that date}.

@@

    \begin{figure}[!ht]
      \centering
      \includegraphics[width=0.8\linewidth]{daily_patterns_}
      \caption{Usage patterns become easy to spot in the requests per hour heatmap}
      \label{fig:dp}
    \end{figure}

>>>>

  The default landing page of the dashboard presents an overview of all the endpoints, 
  which presents some basic utilization information as:
  (1) the number of calls to that endpoint since the begining of tracking
  (2) the number of calls in the last seven days, and
  (3) the number of calls in the last day\footnote{Although very useful, it was only after several months of using the system that the client requested the extra columsn for one day and seven days. An argument for slow research}s
  \va{Please remove the last footnote - it's the equivalent of shooting yourself in the leg.}

@@

        \item \epTranslationsColor is an indicator of the amount of foreign language reading the users are doing.

@@

  For a more detailed view of utilization evolution, \Fref{fig:aeu} shows the \perspective{Daily API Utilization} perspective on API utilization that \tool provides: a stacked bar chart of the number of hits to various endpoints grouped by day\footnote{Endpoint colors are the same in different views}. \Fref{fig:aeu} in particular shows a peak utilization, on June 28, 2017 when the API had more than 12.000 hits\footnote{Turns out that the high school students that were using the system for their French-learning class had a deadline on that date}.

@@

    \begin{figure}
      \centering
      \includegraphics[width=0.8\linewidth]{daily_patterns_}
      \caption{Usage patterns become easy to spot in the requests per hour heatmap}
      \label{fig:dp}
    \end{figure}
